---
title: "p8105_hw1_yl5829"
author: "Yulin Liu"
date: "2024-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown
# Problem 1:

```{r}
#options(repos = c(CRAN = "https://cloud.r-project.org"))
```

```{r}
data("penguins", package = "palmerpenguins")
```

```{r}
#see the row and column components of the data "penguins"
penguins
names(penguins)
```

The data contains 8 columns: Species, island, bill_length_ in mm_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, year.
The species introduce what type of species the penguins is (Adélie, Chinstrap and Gentoo). factor
The island introduce which island the penguin is living on. (Biscoe, Dream or Torgersen). factor
The bill_length_mm denote to bill length in millimeters. numeric
The bill_depth_mm denote to bill depth in millimeters. numeric
The flipper_length_mm denote to an flipper length in millimeters. integer
The body mass denote to the weight of penguin in grams. integer
The sex meaning the sex of penguin (female, male). factor
The year denote the year when the study started (2007, 2008, or 2009). integer


```{r}
#the size of penguins dataset
#number of rows
nrow(penguins)
#number of columns
ncol(penguins)

#the mean flipper length
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

# Problem 2:
```{r}
#random sample of size 10 in standard Normal distribution
random_sample <- rnorm(10)
#a logical vector indicating whether elements of the sample are greater than 0
logical <- random_sample > 0
#a character vector of length 10
cha_vector <-letters[1:10]
#a factor vector of length 10, with 3 different factor “levels”
factor_vector <- factor(rep(c("Level 1", "Level 2", "Level 3"), length.out = 10))
```

```{r}
#combine all the data
combine <- tibble(random_sample, logical, cha_vector, factor_vector)
combine
```

```{r}
#compute the mean
mean(pull(combine, random_sample))
mean(pull(combine, logical))
mean(pull(combine, cha_vector))
mean(pull(combine, factor_vector))
```
It seems like character vector and factor vector are not work while random sampel and logical vector work.

```{r}
#Aplly as.numeric to logical, character, and factor
as.numeric(pull(combine, logical))
as.numeric(pull(combine, cha_vector))
as.numeric(pull(combine, factor_vector))
```
The as.numeric turn the value in logical vector from TURE/FALSE to 1/0, and this is why the mean is working in logical vector as it compute the proportion of TURE
The as.numeric convert NA in character vector because character cannot be directly transfer into numbers, and that's why the mean failed
The as.numeric convert factor into numbers but failed in mean because mean cannot calculate factors.

